---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r options, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# cbRt
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Linux Build Status](https://travis-ci.org/emraher/cbRt.svg?branch=master)](https://travis-ci.org/emraher/cbRt)
[![Windows Build status](https://ci.appveyor.com/api/projects/status/i4g4anmhv22t959x?svg=true)](https://ci.appveyor.com/project/emraher/cbRt)
[![](http://www.r-pkg.org/badges/version/cbRt)](http://www.r-pkg.org/pkg/cbRt)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/cbRt)](http://www.r-pkg.org/pkg/cbRt)
[![Coverage status](https://codecov.io/gh/emraher/cbRt/branch/master/graph/badge.svg)](https://codecov.io/github/emraher/cbRt?branch=master)

## Disclaimer

This software is in no way affiliated, endorsed, or approved by the CBRT. It comes with absolutely no warranty. Also see the [Disclaimer on CBRT webpage](http://www.tcmb.gov.tr/wps/wcm/connect/TCMB+EN/TCMB+EN/Footer+Menu/Disclaimer) since it mentions; **"Information published in this site may be quoted by specific reference thereto, but the use of such information for commercial purposes shall be subject to prior written permission of the CBRT."**

## Overview

`cbRt` is an R interface to The Electronic Data Delivery System (EDDS) of Central Bank of the Republic of Turkey (CBRT). "The system provides a rich range of economic data  and information to support economic education and foster economic research."<sup>[[1]](https://evds2.tcmb.gov.tr/help/videos/hakkindaEN.pdf)</sup>

## Installation

You can install `cbRt` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("emraher/cbRt")
```
## Usage and Arguments

There is only one function in the package which retrieves data from EDDS.

```{r usage, eval = FALSE}
cbRt_get(series, startDate, endDate, token, 
         aggregationTypes = NULL, formulas = NULL, freq = NULL, nd = TRUE,
         as = c("tibble", "tibbletime", "data.frame", "data.table"))
```

Following are the arguments used in the functions. Descriptions are taken from [EDDS Web Service Usage Guide](https://evds2.tcmb.gov.tr/help/videos/EVDS_Web_Service_Usage_Guide.pdf).

**Note that,** `series`**,** `startDate`**,** `endDate`**,** **and** `token` **are all required arguments.**

### `series`
`series` argument can be obtained from CBRT webpage. [Search CBRT webpage](https://evds2.tcmb.gov.tr) in order to find out the series code and use it in the `series` argument. For example, `TP.DK.USD.A` is the code for **(USDTRY) US Dollar (Buying) Exchange Rate**. Argument can take multiple values. 

### `startDate`
`startDate` argument is the series start date as `dd-mm-yyyy` format. In order to display the frequency of the desired series, the first day of the corresponding frequency must be stated in the start date as `dd-mm-yyyy` format. **This argument takes a single value.**

### `endDate`
`endDate` argument is the series end date as `dd-mm-yyyy` format. **This argument takes a single value.**

### `token`
`token` argument is the required API key. [Follow the instructions on CBRT webpage](https://evds2.tcmb.gov.tr/help/videos/EVDS_Web_Service_Usage_Guide.pdf) to obtain the user specific API key. **This argument takes a single value.**

### `aggregationTypes`
`aggregationTypes` argument is the aggregation applied to series. Available aggregations are;

- **default** : Original
- **avg**     : Average
- **min**     : Minimum
- **max**     : Maximum
- **first**   : Beginning
- **last**    : End
- **sum**     : Cumulative

If this parameter is not entered by the user, the original observation parameter is applied for the relevant series. Argument can take multiple values.

### `formulas`
`formulas` argument is the formula applied to series. Available formulas are;

- **0**: Level
- **1**: Percentage Change
- **2**: Difference
- **3**: Year-to-year Percent Change
- **4**: Year-to-year Differences
- **5**: Percentage Change Compared to End-of-Previous Year
- **6**: Difference Compared to End-of-Previous Year
- **7**: Moving Average
- **8**: Moving Sum

If this parameter is not entered by the user, the level formula parameter is applied for the relevant series. Argument can take multiple values.

### `freq`
`freq` argument is the frequency of the series. Available frequencies are;

- **1**: Daily
- **2**: Business
- **3**: Weekly (Friday)
- **4**: Twicemonthly
- **5**: Monthly
- **6**: Quarterly
- **7**: Semiannual
- **8**: Annual

If this parameter is not entered by the user, the common frequency of the series is taken. If you enter a higher frequency (eg: monthly) than the common frequency of the series (eg: annually), the common frequency of the series is taken into account (eg: annually). **This argument takes a single value.** 

### `nd`
`nd` argument is a `TRUE` or `FALSE` argument. Data retrieved sometimes includes ND terms. If `nd` is set to `TRUE` (which is default), all NDs are converted to NAs.

### `as`
`as` argument is the class of the output. Available classes are;

- **tibble** 
- **tibbletime**
- **data.frame**
- **data.table**

**tibble** is the default output class.

## Examples
### Single Series

Following example retrieves **(USDTRY) US Dollar (Buying) Exchange Rate**. Since no optional arguments are supplied, default values for the relevant series are applied. In this case `freq` is `1` (Daily), `aggregationTypes` is `default` (Original), and `formulas` is `0` (Level). Compare the data sets retrieved.

```{r example1}
library(cbRt)
series    <- "TP.DK.USD.A"
startDate <- "01-01-2017"
endDate   <- "01-01-2018"
token     <- readLines("token")

(usd_try <- cbRt_get(series, startDate, endDate, token))
(usd_try <- cbRt_get(series, startDate, endDate, token, aggregationTypes = "default", formulas = 0, freq = 1))
```

We can get single series with multiple different formulas. In this example, `formulas` is set to `0` (Level) and `1` (Percentage Change).

```{r example2}
(usd_try <- cbRt_get(series, startDate, endDate, token, formulas = c(0, 1)))
```

**TODO:** Throw error in this case.

We cannot get single series with multiple different aggregation types, which makes sense. In this example, `aggregationTypes` is set to `avg` (Average) and `min` (Minimum). Result is daily original observations.

```{r example3}
(usd_try <- cbRt_get(series, startDate, endDate, token, aggregationTypes = c("avg", "min")))
```

**TODO:** Gives an extra column. Delete this.

We can get single series with multiple different aggregation types and multiple different formulas. In the first column `aggregationTypes` is `min` (Minimum) and `formulas` is `1` (Percentage Change). In the second column `aggregationTypes` is `max` (Maximum) and `formulas` is `2` (Difference). 

```{r example4}
(usd_try <- cbRt_get(series, startDate, endDate, token, aggregationTypes = c("min", "max"), formulas = c(1, 2)))
```

### Multiple Series

Following example retrieves **(USDTRY) US Dollar (Buying) Exchange Rate** and **(EURTRY) Euro (Buying) Exchange Rate**. Since no optional arguments are supplied, default values for the relevant series are applied. In this case `freq` is `1` (Daily), `aggregationTypes` is `default` (Original), and `formualas` is `0` (Level). Compare the data sets retrieved.

```{r example5}
series <- c("TP.DK.USD.A", "TP.DK.EUR.A")

(usd_eur_try <- cbRt_get(series, startDate, endDate, token))
(usd_eur_try <- cbRt_get(series, startDate, endDate, token, freq = 1, formulas = 0))
```

**TODO:** Gives an extra column. Delete this.

We can get multiple series with multiple different formulas. In this example, `formulas` is set to `0` (Level) and `1` (Percentage Change).

```{r example6}
(usd_eur_try <- cbRt_get(series, startDate, endDate, token, formulas = c(0, 1)))
```

**TODO:** Throw error in this case.

We cannot get multiple series with multiple different aggregation types. In this example, `aggregationTypes` is set to `avg` (Average) and `min` (Minimum). Result is daily original observations.

```{r example7}
(usd_eur_try <- cbRt_get(series, startDate, endDate, token, aggregationTypes = c("avg", "min")))
```

**TODO:** Gives an extra column. Delete this.

We can get multiple series with multiple different aggregation types and multiple different formulas. In the first column `aggregationTypes` is `avg` (Average) and `formulas` is `0` (Level). In the second column `aggregationTypes` is `min` (Minimum) and `formulas` is `1` (Percentage Change). 

```{r example8}
(usd_eur_try <- cbRt_get(series, startDate, endDate, token, 
                        aggregationTypes = c("avg", "min"), formulas = c(0, 1)))
```

We can get multiple series with multiple different aggregation types and multiple different formulas. In the first column `aggregationTypes` is `avg` (Average) and `formulas` is `0` (Level). In the second column `aggregationTypes` is `min` (Minimum) and `formulas` is `1` (Percentage Change). We also set `freq` to `5` (Monthly).

```{r example9}
(usd_eur_try <- cbRt_get(series, startDate, endDate, token, 
                        aggregationTypes = c("avg", "min"), formulas = c(0, 1), freq = 5))
```


### Series with Location Data

Following are examples for data with location attributes. Series is Hedonic House Price Index (HHPI)(2010=100)(CBRT)(Monthly). `TP.HKFE02` and `TP.HKFE03` are price indices codes for Ankara and Istanbul, respectively. This series can be used to draw maps. At this time there is no support in the package though.

```{r example10}
# Single Series
series <- "TP.HKFE02"
(hhpi <- cbRt_get(series, startDate, endDate, token))
```
```{r example11}
# Multiple Series
series <- c("TP.HKFE02", "TP.HKFE03")
(hhpi <- cbRt_get(series, startDate, endDate, token))
```

