---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r options, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# cbRt

<!-- badges: start -->
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check](https://github.com/emraher/cbRt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/emraher/cbRt/actions/workflows/R-CMD-check.yaml)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.18381733.svg)](https://doi.org/10.5281/zenodo.18381733)
<!-- badges: end -->

## Overview

`cbRt` provides an R interface to the Electronic Data Delivery System (EVDS) of the Central Bank of the Republic of Turkey (CBRT). The package enables easy access to economic data including exchange rates, interest rates, and inflation indicators.

**Features:**

- Full EVDS API v3 support with automatic handling of the 150 observation limit
- Intelligent chunking - automatically splits large date ranges into multiple requests
- Hybrid frequency detection - detects series frequency from metadata with fallback
- Multiple output formats - tibble, tsibble, data.frame, or data.table
- Formula transformations - apply calculations at the API level
- Clean data handling - automatic conversion of ND values to NA

## Installation

Install the development version from GitHub:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("emraher/cbRt")
```

## Getting Started

### API Token Setup

You need an API token from the CBRT EVDS system:

1. Visit [https://evds3.tcmb.gov.tr/](https://evds3.tcmb.gov.tr/)
2. Create an account and generate your API key

Store your token securely in your `.Renviron` file:

```r
# Edit your .Renviron file
usethis::edit_r_environ()

# Add this line:
EVDS_TOKEN = "your_api_key_here"
```

Alternatively, set it for the current session:

```r
Sys.setenv(EVDS_TOKEN = "your_api_key_here")
```

### Basic Usage

```{r example-basic, eval = TRUE}
library(cbRt)

# Get USD/TRY exchange rate
usd_try <- cbrt_get(
  series = "TP.DK.USD.A",
  start_date = "01-01-2020",
  end_date = "01-01-2024",
  token = Sys.getenv("EVDS_TOKEN")
)

head(usd_try)
```

### Multiple Series

Retrieve multiple series at once - the package handles different frequencies automatically:

```{r example-multiple, eval = TRUE}
# Get multiple exchange rates
exchange_rates <- cbrt_get(
  series = c("TP.DK.USD.A", "TP.DK.EUR.A", "TP.DK.GBP.A"),
  start_date = "01-01-2023",
  end_date = "01-01-2024",
  token = Sys.getenv("EVDS_TOKEN")
)

head(exchange_rates)
```

### Applying Formulas

```{r example-formulas, eval = TRUE}
# Get year-over-year percent change
usd_try_yoy <- cbrt_get(
  series = "TP.DK.USD.A",
  start_date = "01-01-2020",
  end_date = "01-01-2024",
  formulas = 3,  # Year-to-year percent change
  token = Sys.getenv("EVDS_TOKEN")
)

head(usd_try_yoy)
```

Available formulas: 0 (level), 1 (percentage change), 2 (difference), 3 (year-to-year percent change), 4 (year-to-year differences), 5 (percentage change vs. end-of-previous year), 6 (difference vs. end-of-previous year), 7 (moving average), 8 (moving sum).

### Exploring Available Data

The package includes bundled metadata for all available series, so you can explore data offline:

```{r example-bundled-meta, eval = FALSE}
library(dplyr)
library(stringr)

# Use bundled metadata (no API call needed)
data(cbrt_meta_data)

# Search for exchange rate series
cbrt_meta_data %>%
  filter(str_detect(SERIE_NAME_ENG, regex("exchange", ignore_case = TRUE))) %>%
  select(SERIE_CODE, SERIE_NAME_ENG, FREQUENCY_STR, START_DATE, END_DATE) %>%
  head()
```

Or fetch the latest metadata directly from the API:

```{r example-meta, eval = FALSE}
# Get fresh metadata from API
metadata <- cbrt_meta(token = Sys.getenv("EVDS_TOKEN"))

# Search for specific series
metadata %>%
  filter(str_detect(SERIE_CODE, "USD")) %>%
  select(SERIE_CODE, SERIE_NAME_ENG, FREQUENCY_STR) %>%
  head()
```

## API v3 Support

The package handles the EVDS API v3 observation limit automatically:

### Automatic Chunking

For large date ranges, the package:

1. Detects series frequency from API metadata
2. Splits the date range into chunks of ≤150 observations
3. Fetches data in multiple requests
4. Combines results transparently

```{r example-chunking, eval = TRUE}
# This automatically chunks the request
usd_try_long <- cbrt_get(
  series = "TP.DK.USD.A",
  start_date = "01-01-2015",  # 9+ years of daily data
  end_date = "01-01-2024",
  token = Sys.getenv("EVDS_TOKEN")
)

# You'll see messages like:
# "Series TP.DK.USD.A: Fetching data in 5 chunks to handle 150 observation limit."

nrow(usd_try_long)  # Returns all data, not limited to 150
```

### Frequency Detection

Series frequencies are detected from metadata API. If detection fails, defaults to daily frequency. Supported frequencies: daily, business days, weekly, bi-monthly, monthly, quarterly, semi-annual, yearly.

## Output Formats

Choose your preferred output format:

```{r example-formats, eval = FALSE}
# Tibble (default) - for dplyr workflows
data_tibble <- cbrt_get(series, start_date, end_date, as = "tibble")

# Tsibble - for time series analysis with fable/feasts
data_tsibble <- cbrt_get(series, start_date, end_date, as = "tsibble")

# data.frame - base R compatibility
data_df <- cbrt_get(series, start_date, end_date, as = "data.frame")

# data.table - for data.table workflows
data_dt <- cbrt_get(series, start_date, end_date, as = "data.table")
```

## Documentation

- Getting Started Guide: `vignette("getting-started")`
- Function Reference: `?cbrt_get`, `?cbrt_meta`
- Package Website: [https://eremrah.com/cbRt/](https://eremrah.com/cbRt/)
- EVDS API: [https://evds3.tcmb.gov.tr/](https://evds3.tcmb.gov.tr/)

## Similar Packages

- [etaymaz/CBRT](https://github.com/etaymaz/CBRT) - Alternative CBRT data interface

## Citation

If you use this package in your research, please cite:

```r
citation("cbRt")
```

## Contributing

See the [contributing guide](CONTRIBUTING.md) for details.

- Report bugs or request features: [GitHub Issues](https://github.com/emraher/cbRt/issues)
- Submit pull requests: [GitHub Pull Requests](https://github.com/emraher/cbRt/pulls)

## Disclaimer

This software is in no way affiliated, endorsed, or approved by the Central Bank of the Republic of Turkey (CBRT). It comes with absolutely no warranty.

Please see the [CBRT disclaimer](https://www.tcmb.gov.tr/wps/wcm/connect/TCMB+EN/TCMB+EN/Footer+Menu/Disclaimer) which states: *"Information published in this site may be quoted by specific reference thereto, but the use of such information for commercial purposes shall be subject to prior written permission of the CBRT."*

## License

MIT License - see [LICENSE](LICENSE) file for details.
