---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r options, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# cbRt
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Linux Build Status](https://travis-ci.org/emraher/cbRt.svg?branch=master)](https://travis-ci.org/emraher/cbRt)
[![Windows Build status](https://ci.appveyor.com/api/projects/status/i4g4anmhv22t959x?svg=true)](https://ci.appveyor.com/project/emraher/cbRt)
[![](http://www.r-pkg.org/badges/version/cbRt)](http://www.r-pkg.org/pkg/cbRt)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/cbRt)](http://www.r-pkg.org/pkg/cbRt)
[![Coverage status](https://codecov.io/gh/emraher/cbRt/branch/master/graph/badge.svg)](https://codecov.io/github/emraher/cbRt?branch=master)

## Disclaimer

This software is in no way affiliated, endorsed, or approved by the CBRT. It comes with absolutely no warranty. Also see the [Disclaimer on CBRT webpage](http://www.tcmb.gov.tr/wps/wcm/connect/TCMB+EN/TCMB+EN/Footer+Menu/Disclaimer) since it mentions; **"Information published in this site may be quoted by specific reference thereto, but the use of such information for commercial purposes shall be subject to prior written permission of the CBRT."**

## Overview

`cbRt` is an R interface to The Electronic Data Delivery System (EDDS) of Central Bank of the Republic of Turkey (CBRT). "The system provides a rich range of economic data  and information to support economic education and foster economic research."<sup>[[1]](https://evds2.tcmb.gov.tr/help/videos/hakkindaEN.pdf)</sup>

### Similar Packages

* See [etaymaz/CBRT](https://github.com/etaymaz/CBRT)


## Installation

You can install `cbRt` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("emraher/cbRt")
```
## Usage and Arguments

There are two functions in the package.

`cbrt_meta` retrieves all information for all series.

```{r cbrt_meta_usage, eval = FALSE}
series_meta_info <- cbrt_meta(token = NULL)
```

For `token` argument see the definition below. This returns a tibble. One can search this tibble to find series IDs.


The other function is `cbrt_get` which retrieves data from EDDS.

```{r usage, eval = FALSE}
cbrt_get(series, start_date, end_date, token = NULL, nd = TRUE,
         as = c("tibble", "tsibble", "data.frame", "data.table"))
```

Following are the arguments used in the functions. Descriptions are taken from [EDDS Web Service Usage Guide](https://evds2.tcmb.gov.tr/help/videos/EVDS_Web_Service_Usage_Guide.pdf).

**Note that,** `series`**,** `start_date`**,** `end_date`**,** **and** `token` **are all required arguments.**

### `series`
`series` argument can be obtained from CBRT webpage. [Search CBRT webpage](https://evds2.tcmb.gov.tr) in order to find out the series code and use it in the `series` argument. For example, `TP.DK.USD.A` is the code for **(USDTRY) US Dollar (Buying) Exchange Rate**. Argument can take multiple values. 

### `start_date`
`start_date` argument is the series start date as `dd-mm-yyyy` format. In order to display the frequency of the desired series, the first day of the corresponding frequency must be stated in the start date as `dd-mm-yyyy` format. **This argument takes a single value.**

### `end_date`
`end_date` argument is the series end date as `dd-mm-yyyy` format. **This argument takes a single value.**

### `token`
`token` argument is the required API key. [Follow the instructions on CBRT webpage](https://evds2.tcmb.gov.tr/help/videos/EVDS_Web_Service_Usage_Guide.pdf) to obtain the user specific API key. **This argument takes a single value.** 

You can set your token with 

```
Sys.setenv(EVDS_TOKEN = "Iq83AIL5bss")
```

or you can add it to your `.Renviron` file as

```
EVDS_TOKEN = "Iq83AIL5bss"
```

### `nd`
`nd` argument is a `TRUE` or `FALSE` argument. Data retrieved sometimes includes ND terms. If `nd` is set to `TRUE` (which is default), all NDs are converted to NAs.

### `as`
`as` argument is the class of the output. Available classes are;

- **tibble** 
- **tsibble**
- **data.frame**
- **data.table**

**tibble** is the default output class.

## Examples
### Single Series

Following example retrieves **(USDTRY) US Dollar (Buying) Exchange Rate**. 

```{r example1}
library(cbRt)
series    <- "TP.DK.USD.A"
start_date <- "01-01-2017"
end_date   <- "01-01-2018"
token     <- Sys.getenv("EVDS_TOKEN")

(usd_try <- cbrt_get(series, start_date, end_date, token))
```

### Multiple Series

Following example retrieves **(USDTRY) US Dollar (Buying) Exchange Rate** and **(EURTRY) Euro (Buying) Exchange Rate**. 

```{r example5}
series <- c("TP.DK.USD.A", "TP.DK.EUR.A")

(usd_eur_try <- cbrt_get(series, start_date, end_date, token))
```

### Series with Location Data

Following are examples for data with location attributes. Series is Hedonic House Price Index (HHPI)(2010=100)(CBRT)(Monthly). `TP.HKFE02` and `TP.HKFE03` are price indices codes for Istanbul and Ankara, respectively. This series can be used to draw maps. At this time there is no support in the package though.

```{r example10}
# Single Series
series <- "TP.HKFE02"
(hhpi <- cbrt_get(series, start_date, end_date, token))
```
```{r example11}
# Multiple Series
series <- c("TP.HKFE02", "TP.HKFE03")
(hhpi <- cbrt_get(series, start_date, end_date, token))
```

